Comparing human and model-based forecasts of COVID-19 in Germany and Poland

### 2021-10-13 analysis/analysis.Rmd
update figures and manuscript

  p <- df %>%
    dplyr::filter(target_type == ttype) %>%
    ggplot(aes(y = true_value, x = target_end_date)) + 
    geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = Model), 
                   size = 1.1,
                   alpha = 1,
                   position = position_dodge(width = dodgewidth)) + 
    geom_point(data = trend[target_type == ttype],
               aes(shape = Classification)) +
    geom_line(data = unfiltered_data[target_type == ttype & horizon == 2]) + 
    theme_minimal() + 
    theme(panel.grid.major.x = element_blank()) +
    facet_wrap(~ location_target, scales = "free", ncol = 1) + 
    theme(legend.position = "bottom") + 
    scale_x_date(limits = c(min(limits), max(limits)), 
                 minor_breaks = limits) + 
      scale_y_continuous(labels = scale_fn)
  return(p)
}

@@

p <- ens[, .(value = sum(value), 
             epinow = sum(epinow, na.rm = TRUE), 
             crowd = sum(crowd, na.rm = TRUE)), by = c("variable", "Date")] %>%
  ggplot(aes(y = value, x = Date, color = variable)) + 
  # geom_col(aes(fill = variable), position = position_dodge()) + 
  geom_point() + 
  geom_line() + 
  scale_color_manual(values = colcases) + 
  theme_minimal()

>>>>

  p <- df %>%
    dplyr::filter(target_type == ttype) %>%
    ggplot(aes(y = true_value, x = target_end_date)) + 
    geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = Model), 
                   size = 1.1,
                   alpha = 1,
                   position = position_dodge(width = dodgewidth)) + 
    geom_point(data = trend[target_type == ttype], size = 0.7) +
    geom_line(data = unfiltered_data[target_type == ttype & horizon == 2]) + 
    theme_minimal() + 
    theme(panel.grid.major.x = element_blank()) +
    facet_wrap(~ location_target, scales = "free", ncol = 1) + 
    theme(legend.position = "bottom") + 
    scale_x_date(limits = c(min(limits), max(limits)), 
                 minor_breaks = limits) + 
      scale_y_continuous(labels = scale_fn)
  return(p)
}

@@

targetvars <- data.frame(
  variable = c("GM_inc_case", "GM_inc_death", "PL_inc_case", "PL_inc_death"), 
  target = c("Cases - Germany", "Deaths - Germany", "Cases - Poland", "Deaths - Poland")
)

ens <- ens[, .(value = sum(value), 
             epinow = sum(epinow, na.rm = TRUE), 
             crowd = sum(crowd, na.rm = TRUE)), by = c("variable", "Date")] %>%
  merge(targetvars)

# number of submissions from crowd and epinow2
ens[, .(crowd = sum(crowd) / 4, 
        epinow = sum(epinow) / 4)]

# how many models included in the ensemble? 
ens %>%
  summary()

p <- ens %>%
  ggplot(aes(y = value, x = Date)) + 
  facet_wrap(~ target) + 
  # geom_col(aes(fill = variable), position = position_dodge()) + 
  geom_point(size = 0.7) + 
  geom_line() + 
  scale_color_manual(values = colcases) + 
  theme_minimal() + 
  labs(y = "Number of ensemble members", x = "Date")

### 2021-10-13 manuscript/manuscript.Rmd
update figures and manuscript

## Methods

@@

### Forecast targets and interaction with the German and Polish Forecast Hub

@@

### Crowd forecasts

@@

### Model-based forecasts

@@

### Statistical analysis

@@

Forecasts were analysed using the following scoring metrics: The weighted interval score (WIS) (Johannes Bracher et al. 2021), the absolute error, relative bias, and empirical coverage of the 50% and 90% prediction intervals. The WIS is a proper scoring rule (CITE GNEITING) and can be understood as a generalisation of the absolute error to quantile-based forecasts. It can be decomposed into three separate penalties: forecast spread (i.e. uncertainty of forecasts), over-prediction and under-prediction - lower values are always better. While the over- and under-prediction components of the WIS capture the amount of over-prediction and under-prediction in absolute terms, we also look at a relative tendency to make biased forecasts. The bias metric we use captures how much probability mass of the forecast was above or below the true value and therefore represents a general tendency to over- or under-predict in relative terms. It is bound between -1 (all quantiles of the predictive distribution are below the observed value) and 1 (all quantiles are above the observed value). Empirical coverage is the percentage of observed values that fall inside a given prediction interval (e.g. how many observed values fall inside all 50% prediction intervals). Scoring metrics are explained in more detail in Table \@ref(tab:scoring-metrics) in the SI. All scores were calculated using the scoringutils R package (N. Bosse 2020).

At all stages of the evaluation our forecasts were compared to the median ensemble of all other models submitted to the German and Polish Forecast Hub (Hub ensemble). The ensemble that we call "Hub ensemble" in this paper differs from the official Hub ensemble (here called "XXX") as our forecasts were excluded from it to allow for a more straightforward comparison. To enhance interpretability of scores we mainly report WIS relative to the Hub ensemble in the main text, i.e. we divided all scores by those achieved by the Hub ensemble on the same set of forecasts. In addition to comparing our forecasts against the hub ensemble excluding our models, we also assessed the impact of our forecasts on the performance of the forecasting hub by recalculating separate versions of the Hub ensemble with only some (or all) of our forecasts included.

@@

The renewal model exhibited a noticeable tendency towards over-predicting reported cases across all horizons (as captured by our relative bias metric with values between -1 and 1 and 0 ideally). The crowd forecast tended to over-predict at longer forecast horizons, whereas the Hub ensemble showed no systematic bias (Figure 1X). Regardless of a general relative tendency to over-predict, all forecasting approaches incurred larger absolute penalties from over- than from under-prediction (see decomposition of the WIS into absolute penalties for over-prediction, under-prediction and dispersion in Figure X and Table X), implying that over-prediction (e.g. when missing a peak) tended to be more costly in terms of the WIS.

@@

Of our three models, only the renewal model and the crowd forecast were included in the official Forecast Hub median ensemble (“hub-ensemble-realised”), while the convolution model was never included as it was deemed too similar to the existing renewal model, thus potentially introducing bias. Versions that included either all of our models (“hub-ensemble-with-all”) or only one of them (“hub-ensemble-with-X”) were computed retrospectively.

@@

We also considered the impact of our contributions on the mean, rather than the median ensemble. General trends were similar, with the notable exception of the convolution model, which had a consistently positive impact on the median ensemble, but a mixed and mostly slightly negative impact on the mean ensemble. This may happen if a model is more correct directionally relative to the pre-existing ensemble, but overshoots in absolute terms, thereby moving the ensemble too far. For both the mean and the median ensemble, changes in performance from adding or removing models were of a similar order of magnitude, suggesting that at least in this instance, with a relatively small ensemble size, the median ensemble was not necessarily more ‘robust’ to changes than the mean ensemble. However, the ensemble version with all our forecasts included (“hub-ensemble-with-all”) tended to perform relatively better for the median ensemble than the mean ensemble, suggesting that adding more models may be more beneficial or ‘safer’ for the median than for the mean ensemble as directional errors can more easily cancel out than errors in absolute terms.

@@

>>>>

# Methods

@@

## Forecast targets and interaction with the German and Polish Forecast Hub

@@

## Crowd forecasts

@@

## Model-based forecasts

@@

## Analysis

@@

Forecasts were analysed using the following scoring metrics: The weighted interval score (WIS) (Johannes Bracher et al. 2021), the absolute error, relative bias, and empirical coverage of the 50% and 90% prediction intervals. The WIS is a proper scoring rule (CITE GNEITING) and can be understood as a generalisation of the absolute error to quantile-based forecasts. It can be decomposed into three separate penalties: forecast spread (i.e. uncertainty of forecasts), over-prediction and under-prediction - lower values are always better. While the over- and under-prediction components of the WIS capture the amount of over-prediction and under-prediction in absolute terms, we also look at a relative tendency to make biased forecasts. The bias metric we use captures how much probability mass of the forecast was above or below the true value and therefore represents a general tendency to over- or under-predict in relative terms. It is bound between -1 (all quantiles of the predictive distribution are below the observed value) and 1 (all quantiles are above the observed value). Empirical coverage is the percentage of observed values that fall inside a given prediction interval (e.g. how many observed values fall inside all 50% prediction intervals). Scoring metrics are explained in more detail in Table \@ref(tab:scoring-metrics) in the SI. All scores were calculated using the `scoringutils` R package (N. Bosse 2020).

At all stages of the evaluation our forecasts were compared to the median ensemble of all other models submitted to the German and Polish Forecast Hub ("Hub ensemble"). This "Hub ensemble" was retrospectively computed and excludes all our models. It therefore differs from the official Hub ensemble (here called "hub-ensemble-realised") which included crowd forecasts as well as renewal model forecasts.

The ensemble that we call "Hub ensemble" in this paper differs from the

in that we removed our forecasts to allow for a more straightforward comparison. To enhance interpretability of scores we mainly report WIS relative to the Hub ensemble in the main text, i.e. we divided all scores by those achieved by the Hub ensemble on the same set of forecasts. In addition to comparing our forecasts against the hub ensemble excluding our models, we also assessed the impact of our forecasts on the performance of the forecasting hub by recalculating separate versions of the Hub ensemble with only some (or all) of our forecasts included.

Versions that included either all of our models (“hub-ensemble-with-all”) or only one of them (“hub-ensemble-with-X”) were computed retrospectively.

@@

The renewal model exhibited a noticeable tendency towards over-predicting reported cases across all horizons. The crowd forecast tended to over-predict at longer forecast horizons, whereas the Hub ensemble showed no systematic bias (Figure 1X). Regardless of a general relative tendency to over-predict, all forecasting approaches incurred larger absolute penalties from over- than from under-prediction (see decomposition of the WIS into absolute penalties for over-prediction, under-prediction and dispersion in Figure X and Table X), implying that over-prediction (e.g. when missing a peak) tended to be more costly in terms of the WIS.

@@

Of our three models, only the renewal model and the crowd forecast were included in the official Forecast Hub median ensemble (“hub-ensemble-realised”), while the convolution model was never included as it was deemed too similar to the existing renewal model, thus potentially introducing bias. In the official Hub ensemble, there were on average 7.1 models included (including our own), with a median of 7, a minimum of 4 (on December 28 2020 over the Christmas period) and a maximum of 10. Versions that included either all of our models (“hub-ensemble-with-all”) or only one of them (“hub-ensemble-with-X”) were computed retrospectively. An overview of all models and ensemble verisons is shown in Table \@ref(tab:table-ensemble-versions) in the SI.

@@

We also considered the impact of our contributions on the mean, rather than the median ensemble. 
<!-- In general, the median ensemble tended to perform slightly better than the mean ensemble - needs more analysis -->
General trends were similar, with the notable exception of the convolution model, which had a consistently positive impact on the median ensemble, but a mixed and mostly slightly negative impact on the mean ensemble (REFERENCE FIGURE). This may happen if a model is more correct directionally relative to the pre-existing ensemble, but overshoots in absolute terms, thereby moving the ensemble too far. For both the mean and the median ensemble, changes in performance from adding or removing models were of a similar order of magnitude, suggesting that at least in this instance, with a relatively small ensemble size, the median ensemble was not necessarily more ‘robust’ to changes than the mean ensemble. However, the ensemble version with all our forecasts included (“hub-ensemble-with-all”) tended to perform relatively better for the median ensemble than the mean ensemble, suggesting that adding more models may be more beneficial or ‘safer’ for the median than for the mean ensemble as directional errors can more easily cancel out than errors in absolute terms.

@@

```{r table-ensemble-versions}

rbind(
  list(
    Name = "Hub-ensemble-realised", 
    `Explanation` = "Official Forecast Hub median ensemble. Created by the Forecast Hub officially under the name 'KITCOVIDhub-median\\_ensemble' and used as the default ensemble. Included are our crowd forecasts as well as the renewal model (with one missed submission on December 28 2020, but not the convolution model which was deemed to similar to the renewal model."
  ),
  list(
    Name = "Hub-ensemble-realised-mean", 
    `Explanation` = "Official Forecast Hub mean ensemble. Created by the Forecast Hub officially under the name 'KITCOVIDhub-mean\\_ensemble'."
  ),
  list(Name = "", Explanation = ""),

  list(
    Name = "Hub-ensemble", 
    `Explanation` = "Version of the official Hub median ensemble which excludes all our contributions."
  ),
  list(
    Name = "Hub-ensemble-mean", 
    `Explanation` = "Version of the official Hub mean ensemble which excludes all our contributions."
  ),
  list(
    Name = "Hub-ensemble-with-renewal, 
    Hub-ensemble-with-renewal-mean", 
    `Explanation` = "Versions of the official Hub ensembles which of our contributions includes only the Renewal model."
  ),
  list(
    Name = "Hub-ensemble-with-crowd, 
    Hub-ensemble-with-crowd-mean", 
    `Explanation` = "Versions of the official Hub ensembles which of our contributions includes only the Crowd forecast."
  ),
  list(
    Name = "Hub-ensemble-with-convolution, 
    Hub-ensemble-with-convolution-mean", 
    `Explanation` = "Versions of the official Hub ensembles which of our contributions includes only the Convolution model (which originally was never included in any official Hub ensemble)."
  ),
    list(
    Name = "Hub-ensemble-with-all, 
    Hub-ensemble-with-all-mean", 
    `Explanation` = "Versions of the official Hub ensembles which includes all our contributions. For cases, this is identical to the official Hub ensembles, but for deaths the convolution model was added."
  ),

  list(Name = "", Explanation = ""),
  list(
    Name = "Crowd forecast", 
    `Explanation` = "Submitted to the Forecast Hub as 'epiforecasts-EpiExpert'"
  ),  
  list(
    Name = "Renewal model", 
    `Explanation` = "Submitted to the Forecast Hub as 'epiforecasts-EpiNow2'"
  ),  
  list(
    Name = "Convolution model", 
    `Explanation` = "Submitted to the Forecast Hub as 'epiforecasts-EpiNow2\\_secondary'"
  )
) %>%
  kableExtra::kbl(format = "latex", booktabs = TRUE, 
                  escape = FALSE,
                  caption = "Overview of the models and ensembles used.",
                  longtable = TRUE,
                  linesep = c('\\addlinespace \\addlinespace')) %>%
  kableExtra::column_spec(1, width = "4.5cm") %>%
  kableExtra::column_spec(2, width = "11.0cm") %>%
  # kableExtra::column_spec(3, width = "4.5cm") %>%
  kableExtra::kable_styling(latex_options = c("striped", "repeat_header"))
